Binocular combination in the autonomic nervous system

### 2024-05-09 manuscript.tex
Retrying the previous commit

@@

\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs

@@

  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}

@@

  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}

@@

\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}

@@

\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}

@@

\title{Binocular combination in the autonomic nervous system}
\author{Federico G. Segala\(^1\), Joel T. Martin\(^{1,2}\), Annie Y. Morsi\(^1\), Alex R. Wade\(^{1,3}\) \& Daniel H. Baker\(^{1,3}\)}
\date{}

@@

\(^2\)School of Philosophy, Psychology and Language Sciences, University of Edinburgh, Edinburgh, United Kingdom.

\(^3\)York Biomedical Research Institute, University of York, Heslington, York, United Kingdom.

@@

\hypertarget{abstract-250-words}{%
\section{Abstract (250 words)}\label{abstract-250-words}}

@@

\hypertarget{significance-statement-120-words}{%
\section{Significance statement (120 words)}\label{significance-statement-120-words}}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The autonomic nervous system regulates many involuntary bodily processes, including the constriction and dilation of the pupils in response to light (\protect\hyperlink{ref-McDougal2015}{McDougal and Gamlin, 2015}). The anatomical pathway from the retina to the subcortical nuclei controlling the pupillary light response (PLR) is well established: it includes the Pretectal Olivary nucleus (PON), the Superior Cervical ganglion and the Edinger-Westphal nucleus, which project to the iris sphincter muscles that directly control the pupil size (\protect\hyperlink{ref-McDougal2015}{McDougal and Gamlin, 2015}, \protect\hyperlink{ref-McDougal2008}{2008}; \protect\hyperlink{ref-Wang2015}{Wang and Munoz, 2015}). Evidence of a binocular component to the PLR is shown by the existence of a consensual response of the pupil (the stimulation of one eye will cause constriction of the other eye; \protect\hyperlink{ref-Wyatt1981}{Wyatt and Musselman, 1981}). The anatomical segregation of the subcortical pathway from the rest of the brain means that this binocular combination of signals must occur independently of the cortical processes of binocular integration required for visual perception. Our recent work (\protect\hyperlink{ref-Segala2023}{Segala et al., 2023}) has shown that the algorithm underlying binocular combination of light in the pupil pathway differs from that in the cortex. Here we extend this paradigm to compare binocular combination in different photoreceptor pathways that feed into the autonomic nervous system.

Different classes of retinal photoreceptors, including cones, rods and melanopsin-containing intrinsically photosensitive retinal ganglion cells (ipRGCs), are directly involved in controlling and maintaining the size of the pupils (\protect\hyperlink{ref-Barrionuevo2018}{Barrionuevo et al., 2018}; \protect\hyperlink{ref-Barrionuevo2014}{Barrionuevo et al., 2014}; \protect\hyperlink{ref-Dacey2005}{Dacey et al., 2005}; \protect\hyperlink{ref-Murray2018}{Murray et al., 2018}; \protect\hyperlink{ref-Spitschan2014}{Spitschan et al., 2014}; \protect\hyperlink{ref-Spitschan2019}{Spitschan and Cajochen, 2019}; \protect\hyperlink{ref-Woelders2018}{Woelders et al., 2018}). The cones have been shown to drive the initial rapid constriction of the pupils (\protect\hyperlink{ref-Mathot2018}{Mathôt, 2018}), while the slower and longer activation of the ipRGCs shows their role in maintaining the constriction over a prolonged period of time and in regulating the post-illumination pupillary response (\protect\hyperlink{ref-Markwell2010}{Markwell et al., 2010}; \protect\hyperlink{ref-McDougal2010}{McDougal and Gamlin, 2010}). The ipRGCs are a unique photoreceptor class that has been discovered recently (\protect\hyperlink{ref-Provencio2000}{Provencio et al., 2000}). They express the photopigment melanopsin, and are involved in the regulation of the circadian rhythm (\protect\hyperlink{ref-Panda2002}{Panda et al., 2002}; \protect\hyperlink{ref-Provencio2000}{Provencio et al., 2000}; \protect\hyperlink{ref-Ruby2002}{Ruby et al., 2002}), forming a major input to the PON (\protect\hyperlink{ref-Dacey2003}{Dacey et al., 2003}). The first direct evidence of the involvement of the ipRGCs in the PLR was shown in rats by Lucas et al. (\protect\hyperlink{ref-Lucas2003}{2003}) who genetically deleted the melanopsin gene from rats, resulting in the loss of the intrinsic photosensitivity of the cells and a reduced pupil constriction. This same behaviour was later also observed in primates and humans (\protect\hyperlink{ref-Gamlin2007}{Gamlin et al., 2007}), where it was demonstrated that the PLR continues during light presentation even when cone and rod signalling is blocked, indicating the primary role of the ipRGCs in maintaining pupil constriction over a prolonged time.

A technique known as silent substitution (\protect\hyperlink{ref-Estevez1982}{Estévez and Spekreijse, 1982}) can be used to selectively stimulate specific photoreceptor classes. Silent substitution exploits the fact that each photoreceptor class has a distinct spectral tuning that overlaps with the others. Using a multiprimary system, in which the primaries (i.e.~LEDs) have different spectra, it is possible to target one class of photoreceptors while maintaining the others at a constant activity level, effectively silencing them (\protect\hyperlink{ref-Shapiro1996}{Shapiro et al., 1996}; \protect\hyperlink{ref-Spitschan2018}{Spitschan and Woelders, 2018}, this paper also offers a clear explanation of how to implement silent substitution). Studies using silent substitution have shown that the pupils behave differently depending on which photoreceptor or pathway is being stimulated. For example, pupil responses have been shown to be out of phase when the S cones are stimulated, relative to when the ipRGCs or the L+M cones are stimulated (\protect\hyperlink{ref-Cao2015}{Cao et al., 2015}; \protect\hyperlink{ref-Spitschan2014}{Spitschan et al., 2014}). Additionally, when L and M cones are stimulated individually, there appears to be a paradoxical pupil response: increments in L cone stimulation cause a constriction of the pupil, while increments in M cone stimulation lead to a dilation of the pupil (\protect\hyperlink{ref-Murray2018}{Murray et al., 2018}; \protect\hyperlink{ref-Woelders2018}{Woelders et al., 2018}), perhaps due to activity in the L-M opponent pathway.

Binocular combination has been extensively studied in visual perception, where it is mediated by neurons in primary visual cortex (\protect\hyperlink{ref-Hubel1962}{Hubel and Wiesel, 1962}; \protect\hyperlink{ref-Moradi2009}{Moradi and Heeger, 2009}). For pattern vision, binocular summation occurs at threshold, such that a lower contrast is required to detect a target shown to both eyes than a target shown to only one eye (\protect\hyperlink{ref-Baker2018}{Baker et al., 2018}; \protect\hyperlink{ref-Campbell1965}{Campbell and Green, 1965}; \protect\hyperlink{ref-Legge1984}{Legge, 1984}). At higher contrasts, a process of `ocularity invariance' is observed, in which the response to monocularly- and binocularly-presented patterns is equal (\protect\hyperlink{ref-Baker2007}{Baker et al., 2007}; \protect\hyperlink{ref-Ding2006}{Ding and Sperling, 2006}; \protect\hyperlink{ref-Meese2006}{Meese et al., 2006}; \protect\hyperlink{ref-Moradi2009}{Moradi and Heeger, 2009}). This is thought to be due to a process of interocular suppression that compensates for the additional excitatory drive caused by stimulating two eyes. Our recent work (\protect\hyperlink{ref-Segala2023}{Segala et al., 2023}) has shown that cortical signal combination for luminance flicker is substantially more linear than for spatial patterns, whereas there is evidence of interocular suppression in the pupil pathway.

Most studies that use silent substitution to investigate the PLR stimulate only one eye. If the stimulated eye is pharmacologically dilated to control pupil size, the response from the unstimulated eye (known as the consensual response) can be recorded (e.g. \protect\hyperlink{ref-McDougal2010}{McDougal and Gamlin, 2010}; \protect\hyperlink{ref-Spitschan2014}{Spitschan et al., 2014}). While this approach controls the total amount of light incident at the retina, it does not permit investigation of binocular combination for specific photoreceptor pathways. Here we take a different approach, using a custom binocular multiprimary system (see Figure \ref{fig:spectraplots}a,b) that can modulate each (undilated) eye independently, and a binocular eye tracker to measure the diameter of both pupils. We measure the amplitude of pupil modulation (using Fourier analysis) in response to sinusoidal stimulation at 0.5Hz, for stimuli directed at the luminance pathway, L-M cone pathway, S-cone pathway, and melanopsin pathway (see Figure \ref{fig:spectraplots}e-l). Our key comparisons are between monocular and binocular stimulation, and between monocular stimulation and a dichoptic masking condition in which the two eyes are stimulated at different frequencies (0.5 and 0.4Hz). The results are interpreted using a contemporary model of binocular vision (\protect\hyperlink{ref-Meese2006}{Meese et al., 2006}) implemented within a hierarchical Bayesian framework.

\hypertarget{results}{%
\section{Results}\label{results}}

@@

In contemporary models of binocular signal combination, the amount of binocular facilitation is determined by the magnitude of interocular suppression, with strong suppression reducing facilitation (\protect\hyperlink{ref-Kingdom2015}{Kingdom and Libenson, 2015}). We can estimate the strength of interocular suppression by measuring how much monocular responses are reduced when a dichoptic `mask' is shown to the other eye. In our paradigm, the two components flickered at different frequencies (0.5 and 0.4Hz) so that their responses remained distinct in the Fourier spectrum (e.g. \protect\hyperlink{ref-Busse2009}{Busse et al., 2009}). The yellow diamond symbols in Figure \ref{fig:FirstHarmonicPlots}a-d show the target responses in this condition, and in most cases were weaker than the monocular responses (red circles). The strongest dichoptic masking is found in the L-M condition, where we also observed the binocular suppression effect. Suppression can also be estimated from the responses at 0.4Hz (Figure \ref{fig:FirstHarmonicPlots}f-i). The reduced amplitude in the binocular cross condition (purple squares) relative to the 0.4Hz monocular condition (grey circles), and the progressive decline in amplitude of the dichoptic cross response (yellow diamonds) also differ across photoreceptor conditions, showing similar differences to those observed at 0.5Hz.

To quantify the extent of suppression for each photoreceptor pathway, we fitted each data set using a Bayesian hierarchical implementation of a simple binocular combination model (\protect\hyperlink{ref-Meese2006}{Meese et al., 2006}; \protect\hyperlink{ref-Segala2023}{Segala et al., 2023}). Our primary objective was to compare posterior distributions of the weight of interocular suppression, which are shown in Figure \ref{fig:FirstHarmonicPlots}e. Consistent with our earlier observations, the strongest suppressive weight corresponds to the L-M and Melanopsin conditions, and the weakest suppression corresponds to the luminance and S-cone conditions, with virtually no overlap between the posterior distributions for weak and strong suppression. The model fits were of good quality, as shown by the curves in Figure \ref{fig:FirstHarmonicPlots}a-d,f-i. The fitted model parameters are given in Table \ref{tab:paramtable}.

@@

Finally, we inspected the response phase of our four stimulation conditions, given previous reports that these differ across pathways (\protect\hyperlink{ref-Spitschan2014}{Spitschan et al., 2014}), and may be in antiphase for melanopsin and S-cone signals. Figure \ref{fig:phaseplots} shows the phase angles for the first (a) and second (b) harmonic frequencies for binocular stimulation (monocular stimulation produced very similar results). At the first harmonic, melanopsin and S-cone signals differed in phase by more than 90 degrees, though they were not fully in antiphase. The luminance and L-M responses were approximately in antiphase to each other, however this is likely due to our choice to modulate L+M- in the first half-cycle (corresponding to a luminance increase in the luminance condition), and L-M+ in the second half-cycle (corresponding to a luminance decrease in the luminance condition). Had we reversed this phase arrangement, we would likely have seen a close phase correspondence between the L-M and luminance conditions (another way to think about this is that L-cone decreases and M-cone increases are processed like luminance increases). Phase differences between the luminance, S-cone and melanopsin conditions are likely attributable to different lags in phototransduction at the earliest stage (i.e.~in the retina). At the second harmonic frequency (Figure \ref{fig:phaseplots}b), the luminance condition was again out of phase with the other three, and was approximately in quadrature phase with the L-M condition, and in antiphase with both the S-cone and melanopsin conditions (which were in phase with each other).

@@

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

@@

This is the first study to investigate binocular interactions in the melanopsin pathway directly. A previous meta-analysis (\protect\hyperlink{ref-Spitschan2019}{Spitschan and Cajochen, 2019}) indicated that there may be a substantial binocular facilitation effect in the circadian pathway, as indexed by melatonin suppression (melatonin is a hormone released by the pineal gland; its production is suppressed by exposure to bright light, particularly when the melanopsin-containing ipRGCs are stimulated). In brief, monocular stimulation of the ipRGCs requires up to ten times the signal strength to produce an equivalent effect to binocular stimulation. This superadditive effect implies an absence of interocular suppression, and perhaps the presence of either a highly compressive nonlinearity, or an AND-style neural operation. Our findings here are very different - we do not see substantial binocular facilitation effects in response to melanopsin modulation, and our data indicate that interocular suppression is very strong. Moreover, by measuring the full contrast response function, we can rule out compressive nonlinearities, because the functions accelerate at both the first and second harmonic frequencies. The explanation for this difference could be due to different anatomical pathways. Binocular combination in the circadian system likely takes place in the suprachiasmatic nucleus, whereas in the pupil constriction circuit the Edinger-Westphal nucleus is the most likely site of binocular integration (\protect\hyperlink{ref-Mathot2018}{Mathôt, 2018}). Presumably these anatomical differences, and the practical constraints of the two systems, lead to differences in response.

Our recent psychophysical work (\protect\hyperlink{ref-Baker2024}{Baker et al., 2024}) has looked at binocular interactions in the L-M and S-cone pathways, and compared these to the luminance pathway. Using a contrast discrimination paradigm with spatial modulations of luminance and colour, Baker et al. (\protect\hyperlink{ref-Baker2024}{2024}) found equally strong interocular suppression in all three pathways. This is rather different from our pupillometry results here, which demonstrate weaker suppression in the luminance and S-cone pathways than in the L-M pathway (see Figure \ref{fig:FirstHarmonicPlots}). However, the current experiments involve temporal modulations, which are quite distinct from the spatial modulations used in Baker et al. (\protect\hyperlink{ref-Baker2024}{2024}). Our other recent work (\protect\hyperlink{ref-Segala2023}{Segala et al., 2023}) has shown that temporal luminance modulations involve much weaker interocular suppression in the cortical response than do spatial luminance modulations (\protect\hyperlink{ref-Baker2017}{Baker and Wade, 2017}). These different normalization processes might reflect different priorities for spatial and temporal vision. Spatial vision aims to fuse images to provide binocular single vision, and benefits from `ocularity invariance' (\protect\hyperlink{ref-Baker2007}{Baker et al., 2007}), in which visual appearance is constant when viewed with one eye or two. Temporal vision is critical for motion perception, which can involve alternative binocular computations, such as calculating velocity differences between the eyes (\protect\hyperlink{ref-Kaestner2019}{Kaestner et al., 2019}). Of course the present measurements were of pupil size, which may be subject to different anatomical and functional constraints from those in the cortex. Future research should aim to extend our current findings to perception using psychophysical approaches.

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

@@

\hypertarget{materials-and-methods}{%
\section{Materials and methods}\label{materials-and-methods}}

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

@@

\hypertarget{apparatus-stimuli}{%
\subsection{Apparatus \& stimuli}\label{apparatus-stimuli}}

To present synchronised stimulus modulations independently to each eye, two light engines (SpectraTuneLAB: LEDMOTIVE Technologies, LLC, Barcelona, Spain), each with 10 independently addressable LED colour channels, were integrated into a customised viewing system. The light engines were operated via a Python interface to their REST API (\protect\hyperlink{ref-Martin2022}{Martin et al., 2022}), which supports synchronous launch and playback of spectral sequences prepared in advance and stored in JSON format. When preparing the spectral sequences, the age of participants was used to account for the yellowing of the lenses. We calculated silent substitution solutions using the \emph{PySilSub} toolbox (\protect\hyperlink{ref-Martin2023}{Martin et al., 2023}), using linear algebra. The outputs of the two light engines (see Figure \ref{fig:spectraplots}b,c) were calibrated using an Ocean Optics Jaz spectroradiometer, which was wavelength-calibrated to an Argon lamp and intensity calibrated using a NIST-traceable light source. We used the Stockman and Sharpe (\protect\hyperlink{ref-Stockman2000}{2000}) 10-degree cone fundamentals, and estimates of melanopsin absorbance spectra from CIE S 026 (discussed in \protect\hyperlink{ref-Martin2023}{Martin et al., 2023}) to calculate \(\alpha\)-opic irradiance.

@@

\hypertarget{procedure}{%
\subsection{Procedure}\label{procedure}}

>>>>

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}

@@

\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}

@@

   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}

@@

  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}

@@

\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}

@@

\usepackage{bookmark}

@@

\title{Binocular combination in the autonomic nervous system}
\author{Federico G. Segala\(^1\), Aurelio Bruno\(^{1,2}\), Joel T. Martin\(^{1,3}\), Annie Y. Morsi\(^1\), Alex R. Wade\(^{1,4}\)\\
\& Daniel H. Baker\(^{1,4}\)}
\date{}

@@

\(^2\)School of Psychology and Vision Sciences, Leicester, United Kingdom.

\(^3\)School of Philosophy, Psychology and Language Sciences, University of Edinburgh, Edinburgh, United Kingdom.

\(^4\)York Biomedical Research Institute, University of York, Heslington, York, United Kingdom.

@@

\section{Abstract (250 words)}\label{abstract-250-words}

@@

\section{Significance statement (120 words)}\label{significance-statement-120-words}

\section{Introduction}\label{introduction}

The autonomic nervous system regulates many involuntary bodily processes, including the constriction and dilation of the pupils in response to light (\citeproc{ref-McDougal2015}{McDougal and Gamlin, 2015}). The anatomical pathway from the retina to the subcortical nuclei controlling the pupillary light response (PLR) is well established: it includes the Pretectal Olivary nucleus (PON), the Superior Cervical ganglion and the Edinger-Westphal nucleus, which project to the iris sphincter muscles that directly control the pupil size (\citeproc{ref-McDougal2015}{McDougal and Gamlin, 2015}, \citeproc{ref-McDougal2008}{2008}; \citeproc{ref-Wang2015}{Wang and Munoz, 2015}). Evidence of a binocular component to the PLR is shown by the existence of a consensual response of the pupil (the stimulation of one eye will cause constriction of the other eye; \citeproc{ref-Wyatt1981}{Wyatt and Musselman, 1981}). The anatomical segregation of the subcortical pathway from the rest of the brain means that this binocular combination of signals must occur independently of the cortical processes of binocular integration required for visual perception. Our recent work (\citeproc{ref-Segala2023}{Segala et al., 2023}) has shown that the algorithm underlying binocular combination of light in the pupil pathway differs from that in the cortex. Here we extend this paradigm to compare binocular combination in different photoreceptor pathways that feed into the autonomic nervous system.

Different classes of retinal photoreceptors, including cones, rods and melanopsin-containing intrinsically photosensitive retinal ganglion cells (ipRGCs), are directly involved in controlling and maintaining the size of the pupils (\citeproc{ref-Barrionuevo2018}{Barrionuevo et al., 2018}; \citeproc{ref-Barrionuevo2014}{Barrionuevo et al., 2014}; \citeproc{ref-Dacey2005}{Dacey et al., 2005}; \citeproc{ref-Murray2018}{Murray et al., 2018}; \citeproc{ref-Spitschan2014}{Spitschan et al., 2014}; \citeproc{ref-Spitschan2019}{Spitschan and Cajochen, 2019}; \citeproc{ref-Woelders2018}{Woelders et al., 2018}). The cones have been shown to drive the initial rapid constriction of the pupils (\citeproc{ref-Mathot2018}{Mathôt, 2018}), while the slower and longer activation of the ipRGCs shows their role in maintaining the constriction over a prolonged period of time and in regulating the post-illumination pupillary response (\citeproc{ref-Markwell2010}{Markwell et al., 2010}; \citeproc{ref-McDougal2010}{McDougal and Gamlin, 2010}). The ipRGCs are a unique photoreceptor class that has been discovered recently (\citeproc{ref-Provencio2000}{Provencio et al., 2000}). They express the photopigment melanopsin, and are involved in the regulation of the circadian rhythm (\citeproc{ref-Panda2002}{Panda et al., 2002}; \citeproc{ref-Provencio2000}{Provencio et al., 2000}; \citeproc{ref-Ruby2002}{Ruby et al., 2002}), forming a major input to the PON (\citeproc{ref-Dacey2003}{Dacey et al., 2003}). The first direct evidence of the involvement of the ipRGCs in the PLR was shown in rats by Lucas et al. (\citeproc{ref-Lucas2003}{2003}) who genetically deleted the melanopsin gene from rats, resulting in the loss of the intrinsic photosensitivity of the cells and a reduced pupil constriction. This same behaviour was later also observed in primates and humans (\citeproc{ref-Gamlin2007}{Gamlin et al., 2007}), where it was demonstrated that the PLR continues during light presentation even when cone and rod signalling is blocked, indicating the primary role of the ipRGCs in maintaining pupil constriction over a prolonged time.

A technique known as silent substitution (\citeproc{ref-Estevez1982}{Estévez and Spekreijse, 1982}) can be used to selectively stimulate specific photoreceptor classes. Silent substitution exploits the fact that each photoreceptor class has a distinct spectral tuning that overlaps with the others. Using a multiprimary system, in which the primaries (i.e.~LEDs) have different spectra, it is possible to target one class of photoreceptors while maintaining the others at a constant activity level, effectively silencing them (\citeproc{ref-Shapiro1996}{Shapiro et al., 1996}; \citeproc{ref-Spitschan2018}{Spitschan and Woelders, 2018}, this paper also offers a clear explanation of how to implement silent substitution). Studies using silent substitution have shown that the pupils behave differently depending on which photoreceptor or pathway is being stimulated. For example, pupil responses have been shown to be out of phase when the S cones are stimulated, relative to when the ipRGCs or the L+M cones are stimulated (\citeproc{ref-Cao2015}{Cao et al., 2015}; \citeproc{ref-Spitschan2014}{Spitschan et al., 2014}). Additionally, when L and M cones are stimulated individually, there appears to be a paradoxical pupil response: increments in L cone stimulation cause a constriction of the pupil, while increments in M cone stimulation lead to a dilation of the pupil (\citeproc{ref-Murray2018}{Murray et al., 2018}; \citeproc{ref-Woelders2018}{Woelders et al., 2018}), perhaps due to activity in the L-M opponent pathway.

Binocular combination has been extensively studied in visual perception, where it is mediated by neurons in primary visual cortex (\citeproc{ref-Hubel1962}{Hubel and Wiesel, 1962}; \citeproc{ref-Moradi2009}{Moradi and Heeger, 2009}). For pattern vision, binocular summation occurs at threshold, such that a lower contrast is required to detect a target shown to both eyes than a target shown to only one eye (\citeproc{ref-Baker2018}{Baker et al., 2018}; \citeproc{ref-Campbell1965}{Campbell and Green, 1965}; \citeproc{ref-Legge1984}{Legge, 1984}). At higher contrasts, a process of `ocularity invariance' is observed, in which the response to monocularly- and binocularly-presented patterns is equal (\citeproc{ref-Baker2007}{Baker et al., 2007}; \citeproc{ref-Ding2006}{Ding and Sperling, 2006}; \citeproc{ref-Meese2006}{Meese et al., 2006}; \citeproc{ref-Moradi2009}{Moradi and Heeger, 2009}). This is thought to be due to a process of interocular suppression that compensates for the additional excitatory drive caused by stimulating two eyes. Our recent work (\citeproc{ref-Segala2023}{Segala et al., 2023}) has shown that cortical signal combination for luminance flicker is substantially more linear than for spatial patterns, whereas there is evidence of interocular suppression in the pupil pathway.

Most studies that use silent substitution to investigate the PLR stimulate only one eye. If the stimulated eye is pharmacologically dilated to control pupil size, the response from the unstimulated eye (known as the consensual response) can be recorded (e.g. \citeproc{ref-McDougal2010}{McDougal and Gamlin, 2010}; \citeproc{ref-Spitschan2014}{Spitschan et al., 2014}). While this approach controls the total amount of light incident at the retina, it does not permit investigation of binocular combination for specific photoreceptor pathways. Here we take a different approach, using a custom binocular multiprimary system (see Figure \ref{fig:spectraplots}a,b) that can modulate each (undilated) eye independently, and a binocular eye tracker to measure the diameter of both pupils. We measure the amplitude of pupil modulation (using Fourier analysis) in response to sinusoidal stimulation at 0.5Hz, for stimuli directed at the luminance pathway, L-M cone pathway, S-cone pathway, and melanopsin pathway (see Figure \ref{fig:spectraplots}e-l). Our key comparisons are between monocular and binocular stimulation, and between monocular stimulation and a dichoptic masking condition in which the two eyes are stimulated at different frequencies (0.5 and 0.4Hz). The results are interpreted using a contemporary model of binocular vision (\citeproc{ref-Meese2006}{Meese et al., 2006}) implemented within a hierarchical Bayesian framework.

\section{Results}\label{results}

@@

In contemporary models of binocular signal combination, the amount of binocular facilitation is determined by the magnitude of interocular suppression, with strong suppression reducing facilitation (\citeproc{ref-Kingdom2015}{Kingdom and Libenson, 2015}). We can estimate the strength of interocular suppression by measuring how much monocular responses are reduced when a dichoptic `mask' is shown to the other eye. In our paradigm, the two components flickered at different frequencies (0.5 and 0.4Hz) so that their responses remained distinct in the Fourier spectrum (e.g. \citeproc{ref-Busse2009}{Busse et al., 2009}). The yellow diamond symbols in Figure \ref{fig:FirstHarmonicPlots}a-d show the target responses in this condition, and in most cases were weaker than the monocular responses (red circles). The strongest dichoptic masking is found in the L-M condition, where we also observed the binocular suppression effect. Suppression can also be estimated from the responses at 0.4Hz (Figure \ref{fig:FirstHarmonicPlots}f-i). The reduced amplitude in the binocular cross condition (purple squares) relative to the 0.4Hz monocular condition (grey circles), and the progressive decline in amplitude of the dichoptic cross response (yellow diamonds) also differ across photoreceptor conditions, showing similar differences to those observed at 0.5Hz.

To quantify the extent of suppression for each photoreceptor pathway, we fitted each data set using a Bayesian hierarchical implementation of a simple binocular combination model (\citeproc{ref-Meese2006}{Meese et al., 2006}; \citeproc{ref-Segala2023}{Segala et al., 2023}). Our primary objective was to compare posterior distributions of the weight of interocular suppression, which are shown in Figure \ref{fig:FirstHarmonicPlots}e. Consistent with our earlier observations, the strongest suppressive weight corresponds to the L-M and Melanopsin conditions, and the weakest suppression corresponds to the luminance and S-cone conditions, with virtually no overlap between the posterior distributions for weak and strong suppression. The model fits were of good quality, as shown by the curves in Figure \ref{fig:FirstHarmonicPlots}a-d,f-i. The fitted model parameters are given in Table \ref{tab:paramtable}.

@@

Finally, we inspected the response phase of our four stimulation conditions, given previous reports that these differ across pathways (\citeproc{ref-Spitschan2014}{Spitschan et al., 2014}), and may be in antiphase for melanopsin and S-cone signals. Figure \ref{fig:phaseplots} shows the phase angles for the first (a) and second (b) harmonic frequencies for binocular stimulation (monocular stimulation produced very similar results). At the first harmonic, melanopsin and S-cone signals differed in phase by more than 90 degrees, though they were not fully in antiphase. The luminance and L-M responses were approximately in antiphase to each other, however this is likely due to our choice to modulate L+M- in the first half-cycle (corresponding to a luminance increase in the luminance condition), and L-M+ in the second half-cycle (corresponding to a luminance decrease in the luminance condition). Had we reversed this phase arrangement, we would likely have seen a close phase correspondence between the L-M and luminance conditions (another way to think about this is that L-cone decreases and M-cone increases are processed like luminance increases). Phase differences between the luminance, S-cone and melanopsin conditions are likely attributable to different lags in phototransduction at the earliest stage (i.e.~in the retina). At the second harmonic frequency (Figure \ref{fig:phaseplots}b), the luminance condition was again out of phase with the other three, and was approximately in quadrature phase with the L-M condition, and in antiphase with both the S-cone and melanopsin conditions (which were in phase with each other).

@@

\section{Discussion}\label{discussion}

@@

This is the first study to investigate binocular interactions in the melanopsin pathway directly. A previous meta-analysis (\citeproc{ref-Spitschan2019}{Spitschan and Cajochen, 2019}) indicated that there may be a substantial binocular facilitation effect in the circadian pathway, as indexed by melatonin suppression (melatonin is a hormone released by the pineal gland; its production is suppressed by exposure to bright light, particularly when the melanopsin-containing ipRGCs are stimulated). In brief, monocular stimulation of the ipRGCs requires up to ten times the signal strength to produce an equivalent effect to binocular stimulation. This superadditive effect implies an absence of interocular suppression, and perhaps the presence of either a highly compressive nonlinearity, or an AND-style neural operation. Our findings here are very different - we do not see substantial binocular facilitation effects in response to melanopsin modulation, and our data indicate that interocular suppression is very strong. Moreover, by measuring the full contrast response function, we can rule out compressive nonlinearities, because the functions accelerate at both the first and second harmonic frequencies. The explanation for this difference could be due to different anatomical pathways. Binocular combination in the circadian system likely takes place in the suprachiasmatic nucleus, whereas in the pupil constriction circuit the Edinger-Westphal nucleus is the most likely site of binocular integration (\citeproc{ref-Mathot2018}{Mathôt, 2018}). Presumably these anatomical differences, and the practical constraints of the two systems, lead to differences in response.

Our recent psychophysical work (\citeproc{ref-Baker2024}{Baker et al., 2024}) has looked at binocular interactions in the L-M and S-cone pathways, and compared these to the luminance pathway. Using a contrast discrimination paradigm with spatial modulations of luminance and colour, Baker et al. (\citeproc{ref-Baker2024}{2024}) found equally strong interocular suppression in all three pathways. This is rather different from our pupillometry results here, which demonstrate weaker suppression in the luminance and S-cone pathways than in the L-M pathway (see Figure \ref{fig:FirstHarmonicPlots}). However, the current experiments involve temporal modulations, which are quite distinct from the spatial modulations used in Baker et al. (\citeproc{ref-Baker2024}{2024}). Our other recent work (\citeproc{ref-Segala2023}{Segala et al., 2023}) has shown that temporal luminance modulations involve much weaker interocular suppression in the cortical response than do spatial luminance modulations (\citeproc{ref-Baker2017}{Baker and Wade, 2017}). These different normalization processes might reflect different priorities for spatial and temporal vision. Spatial vision aims to fuse images to provide binocular single vision, and benefits from `ocularity invariance' (\citeproc{ref-Baker2007}{Baker et al., 2007}), in which visual appearance is constant when viewed with one eye or two. Temporal vision is critical for motion perception, which can involve alternative binocular computations, such as calculating velocity differences between the eyes (\citeproc{ref-Kaestner2019}{Kaestner et al., 2019}). Of course the present measurements were of pupil size, which may be subject to different anatomical and functional constraints from those in the cortex. Future research should aim to extend our current findings to perception using psychophysical approaches.

\section{Conclusions}\label{conclusions}

@@

\section{Materials and methods}\label{materials-and-methods}

\subsection{Participants}\label{participants}

@@

\subsection{Apparatus \& stimuli}\label{apparatus-stimuli}

To present synchronised stimulus modulations independently to each eye, two light engines (SpectraTuneLAB: LEDMOTIVE Technologies, LLC, Barcelona, Spain), each with 10 independently addressable LED colour channels, were integrated into a customised viewing system. The light engines were operated via a Python interface to their REST API (\citeproc{ref-Martin2022}{Martin et al., 2022}), which supports synchronous launch and playback of spectral sequences prepared in advance and stored in JSON format. When preparing the spectral sequences, the age of participants was used to account for the yellowing of the lenses. We calculated silent substitution solutions using the \emph{PySilSub} toolbox (\citeproc{ref-Martin2023}{Martin et al., 2023}), using linear algebra. The outputs of the two light engines (see Figure \ref{fig:spectraplots}b,c) were calibrated using an Ocean Optics Jaz spectroradiometer, which was wavelength-calibrated to an Argon lamp and intensity calibrated using a NIST-traceable light source. We used the Stockman and Sharpe (\citeproc{ref-Stockman2000}{2000}) 10-degree cone fundamentals, and estimates of melanopsin absorbance spectra from CIE S 026 (discussed in \citeproc{ref-Martin2023}{Martin et al., 2023}) to calculate \(\alpha\)-opic irradiance.

@@

\subsection{Procedure}\label{procedure}

Before the start of each experiment, participants adjusted the objective planes of the lenses with the help of the experimenter until the stimulus was in focus and they perceived the two pieces of blackout material as one fused disc. Pupil responses to binocular temporal contrast modulations were examined in a factorial design that combined six ocular conditions and five temporal contrast levels: 6, 12, 24, 48 and 96\% of the available dynamic range. This design, similar to that used in our previous studies (\citeproc{ref-Baker2020}{Baker et al., 2020}; \citeproc{ref-Baker2017}{Baker and Wade, 2017}; \citeproc{ref-Segala2023}{Segala et al., 2023}), was applied in four separate experiments, each with a different mode of photoreceptor stimulation. In the first three conditions, the discs flickered at 0.5 Hz, in either a monocular, binocular or dichoptic arrangement. In the dichoptic condition the non-target eye saw a flickering fixed contrast of 48\% of the available dynamic range. In the remaining three conditions (the cross-frequency conditions) one eye's disc flickered at 0.4 Hz, and the other eye's disc flickered at 0.5 Hz. This included monocular responses at 0.4 Hz, as well as binocular (one eye sees each frequency at the target contrast) and dichoptic (target stimulus flickering at 0.5 Hz, mask contrast of 48\% at 0.4 Hz in the other eye) arrangements. We counterbalanced presentation of the target stimulus across the left and right eyes.
